### 2022-02-28 introduction/introduction.tex
wan main comments

  After spending so much time and energy on developing higher-order provers, we
  took a step back and worked on improving first-order provers. As they form the
  basis of higher-order reasoning in our approach, this should improve
  higher-order performance as well. In particular, we looked at the most
  efficient approaches to simplify problems in propositional logic, which is
  simpler than first-order logic and even decidable. We lifted some of these
  propositional approaches to first-order logic and implemented them in Zipperposition.

>>>>

  After spending so much time and energy on developing higher-order provers, we
  took a step back and worked on improving first-order provers. As they form the
  basis of higher-order reasoning in our approach, this should improve
  higher-order performance as well. In particular, we looked at
  efficient approaches to simplify problems in propositional logic, which is
  simpler than first-order logic and even decidable. We lifted some of these
  propositional approaches to first-order logic and implemented them in Zipperposition.

### 2022-02-28 pre/pre.tex
wan main comments

We assume the standard notions of $\alpha$, $\beta$ and $\eta$ conversions and
write $s \equi t$ if terms $s$ and $t$ are equal modulo $\alpha\beta\eta$
converison. We let $s \, \overline{t}_n$ stand for $s \, t_1 \, \ldots \, t_n$
and $\lam{\overline{x}_n}{s}$ for $\lambda x_1. \ldots \lambda x_n. \> s$. Every
$\beta$-reduced term can be written as $\lam{\tuple{x}{m}}{a \, \tuplen{t}}$,
where $a$ is not an application; we call $a$ the \emph{head} of the term. By
convention, $a, b$ or $\zeta$ denote heads. If $a$ is a free variable, we call the term
\emph{flex}; otherwise, the term is \emph{rigid}.

@@

Higher-order substitutions ($\sigma,\varrho,\ldots$) are functions from free and bound
variables to terms. A variable $F$ is mapped by $\sigma$ if $ \sigma(F) \not\equi
F$. Each substitution maps only finitely many variables, and it is denoted as in
the first-order case. Application of  $\sigma$ to term $t$ is denoted $\sigma(t)$;
this application $\alpha$-renames $t$ to avoid variable capture. For example,
$\{X \mapsto x\}(\lamx{X \, \cst{a}})$ results in $\lam{y}{x \, \cst{a}}$.  Given a substitution
$\varrho$, which maps $F$ to $s$, we write $\varrho\setminus\{F \mapsto s\}$ to
denote a substitution that does not map $F$ and otherwise coincides with
$\varrho$. Given substitutions $\varrho$ and $\sigma$, mapping disjoint variable
sets, we write $\varrho \cup \sigma$ to denote $\varrho\sigma$.

@@

In standard, non-clausal first-order logic, an atom is either
predicate or equational. For uniformity, and to
stay close to the implementation, we encode predicate atoms as equations with
$\itrue$. Negative literals are encoded as disequations. For example,
$\cst{even}(x)$ is encoded as $\poslit{\cst{even}(x)}$, and
$\neg\,\cst{even}(x)$ is encoded as $\neglit{\cst{even}(x)}$.

@@

>>>>

\looseness=-1
With $\alpha$-renaming we assume the rule $(\lambda x. \, s) \longrightarrow \lambda
y. \, \{x \mapsto y\}(s)$ where $y$ does not occur loosely bound in $s$ and 
$y$ is not bound in $s$. Application of substitution $\sigma$ to term
$t$, implicitly $\alpha$-renames bound variables to avoid capture.  For example,
$\{X \mapsto x\}(\lamx{X \, \cst{a}})$ results in $\lam{y}{x \, \cst{a}}$. The
$\beta$-reduction rule corresponds to applying arguments to function: $(\lambda
x. s) \, t \longrightarrow \{x \mapsto t\} (s)$, where the bound variables in $s$
are renamed to avoid capture. Lastly, $\eta$-reduction is defined as $\lambda
x.\, s \, x \longrightarrow s$ under the condition that $x$ is not loosely
bound in $x$. We write $s \equi t$ if terms $s$ and $t$ are equal modulo $\alpha\beta\eta$
converison.

\looseness=-1
As hinted in previous definitions, higher-order substitutions
($\sigma,\varrho,\ldots$) are functions from both free and bound variables to
terms. A variable $F$ is mapped by $\sigma$ if $ \sigma(F) \not\equi F$.
Substitutions map only finitely many variables. Given a substitution $\varrho$,
which maps $F$ to $s$, with $\varrho\setminus\{F \mapsto s\}$ we denote a
substitution that does not map $F$ and otherwise coincides with $\varrho$. Given
substitutions $\varrho$ and $\sigma$, mapping disjoint variable sets, we write
$\varrho \cup \sigma$ to denote $\varrho\sigma$.

We let $s \, \overline{t}_n$ stand for $s \, t_1 \, \ldots \, t_n$
and $\lam{\overline{x}_n}{s}$ for $\lambda x_1. \ldots \lambda x_n. \> s$. Every
$\beta$-reduced term can be written as $\lam{\tuple{x}{m}}{a \, \tuplen{t}}$,
where $a$ is not an application; we call $a$ the \emph{head} of the term. By
convention, $a, b$ or $\zeta$ denote heads. If $a$ is a free variable, we call the term
\emph{flex}; otherwise, the term is \emph{rigid}.

@@

@@

In standard, non-clausal first-order logic, an atom is either
predicate or equational. For uniformity, and to
stay close to the implementation, we encode predicate atoms as equations with
$\itrue$. Negative literals are encoded as disequations. For example,
$\cst{even}(x)$ is encoded as $\poslit{\cst{even}(x)}$, and
$\neg\,\cst{even}(x)$ is encoded as $\neglit{\cst{even}(x)}$. To lighten the notation,
we sometimes write the predicate literals in non-encoded form.

@@

\medskip
Let us give some intuition behind how these rules are applied in practice. The
\infname{ER} rule is used to replace equality with equality within a clause.
Consider the formula $\iforall x.\,  x \eq \cst{2} \iimplies \cst{even}(x)$. In
clausal form it is $X \noteq \cst{2} \llor \cst{even}(X)$. Applying the
\infname{ER} rule effectively applies the precondition $X \eq \cst{2}$ to the
conclusion $\cst{even}(X)$ to obtain $\cst{even}(\cst{2})$. The \infname{EF}
rule can be used to find instances of the clause that can further be simplified.
Consider the clause $\cst{f}(X) \eq \cst{a} \llor \cst{f}(Y) \eq \cst{a} \llor
\cst{p}(X, Y)$. Applying \infname{EF} to its first two literals yields $\cst{a}
\noteq \cst{a} \llor \cst{f}(X) \eq \cst{a} \llor \cst{p}(X, X)$. This clause
then further simplifies to $\cst{f}(X) \eq \cst{a} \llor \cst{p}(X, X)$ using \infname{ER}.

Lastly, superposition rules are used to simulate what mathematicians do when
they ``replace equals by equals''. Left premise intuitively states that an
equation $s \eq t$ holds under condition $C$ (or more precisely, its negation).
Similarly, the right premise asserts that a (dis)equation holds under a
condition. Superposition rule simply concatenates the conditions for both
premises and replaces equals by equals in two main inference terms. For example
given clauses $\neg \cst{even}(X) \llor \cst{f}(X) \eq \cst{a}$ and $\neg
\cst{even}(\cst{b}) \llor \cst{g}(\cst{f}(Y)) \eq \cst{b}$, superposition between their
last literals results in $\neg
\cst{even}(X) \llor \neg \cst{even}(\cst{b}) \llor \cst{g}(\cst{a}) \eq
\cst{b}$.
